LSTM for Sentiment Analysis on Twitter

### 2015-12-20 paper/paper.tex
update: conclusion, merge in Jun paper

\documentclass{article} % For LaTeX2e
\usepackage{nips15submit_e,times}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{url}
%\documentstyle[nips14submit_09,times,art10]{article} % For LaTeX 2.09
\newcommand\BibTeX{B{\sc ib}\TeX}

@@

\section{Introduction}

We use LSTM \cite{hochreiter1997long} to do stuff

\section{Related Work}

Dynamic convolutional neural net for sentences \cite{kalchbrenner2014convolutional}

sentiment specific word embeddings \cite{tang2014learning}

character-lvl CNN for text classification \cite{zhang2015character}

\section{Model}

@@

We compare word-level models under two settings: initializing the word embeddings using random vectors (``word/rnd'') and initialization using the embeddings provided in \cite{tang2014learning} (``word/sswe'').

We train all models using the implementation of the Adam algorithm \cite{kigma2014adam} provided in Lasagne \cite{lasagne} and a learning rate set to $0.1$ \footnote{We retrained the ascii/rnd/200 on SemEval using AdaGrad and a learning rate of $0.01$ to achieve $84.13$; using Adam and $0.1$ learning rate, the result was $83.21$}.

\begin{table}[h!]
\centering
\caption{Accuracy across LSTMs}
\label{char-models}
\begin{tabular}{|l|ll|}
\hline
              & 1.6M (acc)          & semeval (acc) \\
\hline
ascii/rnd/50  & \textbf{83.84} & 82.08   \\
ascii/rnd/200 & 82.45          & \textbf{84.13}   \\
ascii/eye/50  & 77.44          & 79.18   \\
utf8          & 81.34          & 82.34   \\
word/rnd      & 81.85          & 78.07   \\
word/sswe     & 83.24          & 79.27   \\
\hline
\end{tabular}
\end{table}

The character-level models using the ``ascii'' character set outperformed the other models. It is worth noting that the ``utf8'' model performaed comparably despite having a much larger vocabulary size. It is not suprising that the ``utf8'' model performed worse than the ``ascii'' model, as the test data consisted of only English tweets; however, since the ``utf8'' model is implicitly multi-lingual, our results suggest that the same model may perform well across multiple languages. We leave such experiments for future work.

\subsection{Analysis}
Figure ~\ref{fig:cool} shows the effect of character repetition on model confidence over the course of the sequence, where confidence is computed using the softmax function:

@@

\begin{figure}[h!]
\label{fig:cool}
\begin{center}
%\framebox[4.0in]{$\;$}
\includegraphics[width=0.95\textwidth]{figs/cool}
%\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
\end{center}
\caption{Sample figure caption.}
\end{figure}

@@

\begin{figure}[h!]
\label{fig:puppies}
\begin{center}
%\framebox[4.0in]{$\;$}
\includegraphics[width=0.95\textwidth]{figs/puppies}
%\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
\end{center}
\caption{Sample figure caption.}
\end{figure}

@@

\begin{figure}[h!]
\label{fig:dentist}
\begin{center}
%\framebox[4.0in]{$\;$}
\includegraphics[width=0.95\textwidth]{figs/dentist}
%\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
\end{center}
\caption{Confidence in ``negative'' prediction over time}
\end{figure}

\section{Conclusion}

\subsubsection*{Acknowledgments}

>>>>

\documentclass{article} % For LaTeX2e
\usepackage{nips15submit_e,times}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{url}
\usepackage{amsmath}
%\documentstyle[nips14submit_09,times,art10]{article} % For LaTeX 2.09
\newcommand\BibTeX{B{\sc ib}\TeX}

@@

\section{Introduction}
% why analysis sentiment expecially twitter
In the last few years, microblogging has become a very popular communication tool in people's social life.
On microblogging platforms, such as Twitter\footnote{\url{https://twitter.com/}} and Facebook\footnote{\url{https://www.facebook.com/}}, a diverse range of people are attracted to post short sentences, images, and video links to share life issues and opinions.
This popularity results in enormous amount of information covering a wide range of topics on the brands, products, politics and social events. 
Such data is a valuable and efficient source for marketing and social studies. 
Sentiment analysis on microblogging has obtained special interest, because it determines the attitude of a user with respect to some topic or product and thus provides provide convincing information
For example, it may help manufacturing companies to know how people like their product (or service), what would people prefer.

In this paper, we focus on using sentiment analysis on Twitter. 
Twitter is an extremely popular microblogging platform which allows people to post messages of up to 140 characters.
Because of the short nature of tweets, people often post twitter messages (called Tweets) frequently while attending events like product launches, movie premiers, music concerts or just to express their opinion on a trending or current topic.
As such, they can be a valuable source of public opinion or feedback \cite{o2010tweets, bollen2011twitter, bollen2009modeling}.

Realizing this importance, analyzing sentiment of Tweets has been a recurring task in the SemEval \footnote{\url{http://alt.qcri.org/semeval2016/task4/}} competitions.

%We restrict our interest on Twitter for the following reasons: 
%People are posting on various topics on Twitter, thus it is a valuable source of data; Twitter is one of the most popular microblogging platform and the number of messages are growing everyday. The collected corpus can be arbitrary large. Users of Twitter have various background and have different language conventions. It is possible to collect tweets of wide diversity in both topics and language models.

Although there exist plenty of work on text classification, some unique characteristics of tweets present special challenges for sentiment analysis:
1. Tweets are short in length. There is a limitation of $140$ words for each tweet; 2. The language used in tweets is very informal with misspelling, creative spelling,  new words, slangs, and URLs; 3. Emotions and hashtags are frequently used.

% what we did
In this paper, we propose a bi-directional Long Short-Term Memory (LSTM) method for sentiment analysis. We tried both word-level and character-level features on the bi-directional LSTM model and compare the results with Dynamic Convolutional Neural Network (DCNN) \cite{kalchbrenner2014convolutional}.
We show that the accuracy of sentiment analysis ... (Please revise this part.)

We train our model on $1.6$ M distantly-supervised tweets collected by Go et. al. \cite{go2009twitter}, and evaluate the results on SemEval-2016 Task 4\footnote{\url{http://alt.qcri.org/semeval2016/task4/}}. Currently we focus on polarity classification, and plan to apply our model on 5-point scale classification in the future.

% the structure of the paper
In the remaining of this paper, we first introduce the related work, the dataset, and the evaluation methodology. We then describe the model we proposed. Finally, we discuss the experiment results and point out possible directions for future research. (Please revise this part.)

\section{Related Work}
Twitter sentiment analysis is increasingly drawing attention of researchers in recent years. 
Given the length limitations on tweets, sentiment analysis of tweets is often considered similar to sentence-level sentiment analysis \cite{kouloumpis2011twitter}.
However, phrase and sentence level approaches can hardly define the sentiment of some specific topics. Considering opinions adhering on different topics, Wang et. al. \cite{wang2011topic} proposed a hashtag-level sentiment classification method  to generate the overall sentiment polarity for a given hashtag.
Recently, following the work of \cite{mikolov2013efficient} some researchers used neural network to implement sentiment classification. 
For example, Kim \cite{kim2014convolutional} adopted convolutional neural networks to learn sentiment-bearing sentence vectors, Mikolov et al. \cite{mikolov2013distributed} proposed Paragraph vector which outperformed bag-of-words model for sentiment analysis, and Tang et. al. \cite{tang2014learning} used ConvNets to learn sentiment specific word embedding (SSWE), which encodes sentiment information in the continuous
representation of words.
Furthermore, Kalchbrenner \cite{kalchbrenner2014convolutional} proposed a Dynamic Convolutional Neural Network (DCNN) which uses dynamic k-max pooling, a global pooling operation over linear sequences.
Instead of directly applying ConvNets to embeddings of words, \cite{zhang2015character} applies the network only on characters. They showed that the deep ConvNets does not require knowledge of words and thus can work for different languages.
LSTM \cite{hochreiter1997long} is another state-of-the-art semantic composition models for sentiment classification \cite{li2015tree}. Similar to DCNN, it also learns fixed-length vectors for sentences of varying length, captures words order in a sentence and does not depend on external dependency or constituency parse results.

\subsection{Dynamic Convolutional Neural Networks (DCNN)}
We briefly review the architecture of DCNN \cite{kalchbrenner2014convolutional} which has shown state of the art performance for sentiment classification on Twitter. The winning entry for SemEval15 \cite{severynunitn} task on Twitter sentiment classification also used DCNN. Figure \ref{fig:dcnn} summarizes the architecture.

When used for sentiment classification on Twitter, the input to the DCNN is a matrix of word embeddings for each word in the tweet. For example, if the tweet consists of $s$ words then the input to the DCCNN is:
\begin{flalign*}
	S = 
	\begin{bmatrix}
	| & | & \ldots & | \\
	w_1 & w_2 & \ldots & w_s \\
	| & | & \ldots & | \\
	\end{bmatrix}_{k\times s}
\end{flalign*}
where each $w_i \in R^k$ is a $k$-dimensional dense word embedding \cite{mikolov2013distributed}.
The architecture consists of multiple layers of convolutions and max-pooling on top of the input matrix, followed a fully connected layer which is input to a softmax. The convolutions are of type \emph{wide-convolutions} of one-dimension. For example, for the input matrix $S \in R^{k\times s}$, a wide-convolution filter operating on $S$ will consist of convolution weights $m\in R^{k\times c}$ and will result in a matrix having dimension $k \times (s+c-1)$. Note here $c$ is the convolution filter width, which is a hyperparameter.
The max-pooling operations presented in \cite{kalchbrenner2014convolutional} are different from the regular max-pooling. They present {\it dynamic $k$-max pooling}. $k$-max pooling takes the top $k$ maximum activations as opposed to just the maximum activation and the value of $k$ is selected dynamically based on the following formula: $k_l = \max \left( k_{top}, 	\lceil \frac{L-l}{L} s\rceil \right)$, where $l$ is number of current convolution layer, $L$ is total number of convolutions and $k_{top}$ is a fixed hyperparameter. 
Note that while \cite{kalchbrenner2014convolutional} used multiple layers of convolutions and max-pooling, subsequent work found that using a single layer of convolution and max-pooling gives similar results \cite{kim2014convolutional} \cite{severynunitn}.

\begin{figure}
	\centering
	\includegraphics[height=4in, width=0.8\textwidth]{figs/DCNN.png}
	\caption{Dynamic Convulational Neural Network of \cite{kalchbrenner2014convolutional} (Source: \cite{kalchbrenner2014convolutional})}
	\label{fig:dcnn}
\end{figure}

\subsection{Recurrent Neural Networks with Long Short Term Memory (LSTM)}
Recurrent Neural Networks (RNNs) are a class of artificial neural networks used for modeling sequences.
RNNs are highly flexible in their use of context information as they can learn what part of the input sequence to store to memory and what parts to ignore. They also allow modeling of various regimes of sequence modeling as shown in Figure \ref{fig:rnn}. Please refer to \cite{graves2012supervised} for a comprehensive review of sequence modeling using RNN.

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{figs/rnn.jpeg}
	\caption{RNNs allow modeling of multiple types of input and output sequences (Source: \cite{karpathy-rnn})}
	\label{fig:rnn}
\end{figure}

One of the short comings of RNN is that it is very difficult to store information over long sequences because of problems due to vanishing and exploding gradients as explained in \cite{hochreiter2001gradient}.
{\it Long Short-Term Memory (LSTM)} \cite{hochreiter1997long} are designed to remedy this and store information over larger input sequences.
They achieve this using special ``memory cell'' units. 
Figure \ref{fig:lstm} shows the architecture of this cell which consists of an input gate, a forget gate, an output gate and a recurring cell state. 
Refer to \cite{colah} for a gentle introduction to LSTM and to \cite{graves2012supervised} for a more comprehensive review and applications.

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{figs/LSTM.png}
	\caption{The repeating module of LSTM. $x_t$ is the input as time $t$ and $h_t$ is the output from the LSTM output gate at time $t$. The top horizontal line corresponds to the cell state and the bottom line corresponds to the hidden state (both of which are recurring states).
	(Source: \cite{colah})}
	\label{fig:lstm}
\end{figure}

@@

We compare word-level models under two settings: initializing the word embeddings using random vectors (``word/rnd'') and initialization using ``sentiment-specific'' word embeddings provided in \cite{tang2014learning} (``word/sswe'').

Our results are shown in Table ~\ref{table:results}. We train all models using the implementation of the Adam algorithm \cite{kingma2014adam} provided in Lasagne\footnote{\url{https://github.com/Lasagne/Lasagne}} and a learning rate set to $0.1$ \footnote{We retrained the ascii/rnd/200 on SemEval using AdaGrad and a learning rate of $0.01$ to achieve $84.13$; using Adam and $0.1$ learning rate, the result was $83.21$}. To learn word and character embeddings, we use a bidirectional LSTM with $256$ hidden units, followed by a mean pooling and a dropout layer ($p=0.5$), a second forward-directional LSTM (again with $256$ hidden units) and a final dropout layer ($p=0.6$). We obtain final predictions using the softmax function.

\begin{table}[h!]
\label{table:results}
\centering
\caption{Accuracy across LSTMs}
\begin{tabular}{|l|ll|}
\hline
              & 1.6M (acc)          & semeval (acc) \\
\hline
ascii/rnd/50  & 83.84          & 82.08            \\
ascii/rnd/200 & 82.45          & \textbf{84.13}   \\
ascii/eye/50  & 77.44          & 79.18   \\
utf8          & 81.34          & 82.34   \\
char-dcnn     & 75.0           & 81.3    \\
word/rnd      & 81.85          & 78.07   \\
word/sswe     & 83.24          & 79.27   \\
word/dcnn     & \textbf{87.4}\footnote{Reported in \cite{kalchbrenner2014convolutional}}           & 81.3    \\
\hline
\end{tabular}
\end{table}

The character-level models using the ``ascii'' character set outperformed the other models on the SemEval dataset. Due to the highly-productive nature of the Twitter ``lexicon'', users' predisposition toward using slang dialects, and the constraint of the $140$-character limit, it makes sense that word-level models underperform.

It is worth noting that the ``utf8'' model performed comparably despite having a much larger vocabulary size. It is not suprising that the ``utf8'' model performed worse than the ``ascii'' model, as the test data consisted of only English tweets; however, since the ``utf8'' model is implicitly multi-lingual, our results suggest that the same model may perform well across multiple languages. We leave such experiments for future work.

As Table ~\ref{table:results} shows, \cite{kalchbrenner2014convolutional} outperform our character LSTM on the Go dataset. However, the results presented for our character and word-level models achieve competitive performance on the Go dataset without tuning hyperparameters such as learning rate and network width.

\subsection{Qualitative Analysis}
Figure ~\ref{fig:cool} shows the effect of character repetition on model confidence over the course of the sequence, where confidence is computed using the softmax function:

@@

\begin{figure}[h!]
\label{fig:cool}
\begin{center}
%\framebox[4.0in]{$\;$}
\includegraphics[width=0.95\textwidth]{figs/cool}
%\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
\end{center}
\caption{Comparison of model confidence for different forms of the word ``cool''}
\end{figure}

@@

\begin{figure}[h!]
\label{fig:puppies}
\begin{center}
%\framebox[4.0in]{$\;$}
\includegraphics[width=0.95\textwidth]{figs/puppies}
%\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
\end{center}
\caption{Comparison of model confidence for ``I love puppies'' vs. ``I hate puppies''}
\end{figure}

@@

\begin{figure}[h!]
\label{fig:dentist}
\begin{center}
%\framebox[4.0in]{$\;$}
\includegraphics[width=0.95\textwidth]{figs/dentist}
%\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
\end{center}
\caption{Confidence in ``negative'' prediction over time across four tweets}
\end{figure}

\section{Conclusion}
In this paper, we show promising empirical evidence that character-level models perform well for sentiment analysis on tweets, which are ``noisy'', in that users think of millions of new words and spellings of words every day, and ``brief'', in that they are constrained to $140$ characters. This type of text data is becoming an important research focus in the field of NLP, as data is often cheap to collect in high volume and can provide important insight into society-wide trends. Though the experiments in this paper focus on sentiment classification, we believe our results provide a basis for future work on character-level modelling for a variety of other NLP tasks.
